### 2022-07-24 main/paper2.tex
Add files via upload

\section{Introduction}\label{S:1}

@@

In the great majority of practical applications, Granger-causality has been studied in the context of Vector Autoregressive (VAR) models. 
For instance, \citet{billio2012econometric} proposes several measures  based on Granger-causality to capture the connections between the monthly returns of different financial institutions. 
It uses principle component analysis and ``pairwise'' Granger-causality tests to identify the causal networks. 
Other related works are \citet{diebold2014network} and \citet{barigozzi2016network} in which the authors propose connectedness measures based on generalized variance decomposition.
However, the measures introduced in these works are again limited to linear systems and they are based on pairwise comparison which as we show in Section  \ref{sec:dig} fails to infer the true causal relationships.
%and they are based on pairwise comparison which as we show in Section \ref{sec:dig}, it fail to infer the true causal relationships.
 %focuses on one particular network structure: the long-run variance decomposition network. Such networks are characterized by the infinite vector moving average and similar to \citet{diebold2014network}, they are limited to linear systems.

Contributions of this paper are both in network identification literature as well in finance.
%This paper adds contribution in the network identification literature as well as in finance literature. 
Our contribution to network identification are as follows:
\begin{itemize}
    \item We use an information-theoretic measure known as directed information (DI) to infer the Granger-causalities among a set of time series. 
This measure is non-parametric, i.e., it does not depend on the underlying model of the dynamics and it is capable of capturing causal relationships in both linear and nonlinear systems. The output of this approach is a directed graph known as Directed Information Graph (DIG) that visualizes the interconnections among a set of time series such as stock returns.

\item Computing DI has both high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
To overcome this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings. 
\end{itemize}

@@

Applications of DIG are various in finance. 
In particular, we recommend to use it as a preliminary feature selection of another predictive model. Feature selection is a process often used in machine learning and statistics which consists of keeping only a subset of relevant features, usually to avoid overfitting or to reduce dimensionality. 
For example, \citet{fs1} and \citet{fs2} show that extraneous features are prone to reduce model's performance measures. 
Finance applications of feature selection models are various and include credit scoring, stock market behavior analysis  or even fraud detection (\citet{fs_fraud}).
\citet{tsai2009} states that feature selection preprocessing is not addressed carefully enough in the bankruptcy prediction literature. They compare five feature selection methods used in bankruptcy prediction: t-test, correlation matrix, step-wise regression, Principle Component Analysis (PCA) and factor analysis and show that any of these methods improves performance. 
\citet{fs_stock} uses a Sequential Forward Selection algorithm to select relevant features predicting the Turkish market index. The use of such feature selection model reduces model prediction error compared to the case where all features are used. This is due to information embedded in several economic factors already included in the market index. They show that only the lagged value of the market index is enough to predict the forthcoming value of the index.

@@

\subsection{Related Work}
In recent years, several approaches have been developed to generalize the applicability of Granger-causality to non-linear and large dynamics. 
To mention a few, \citet{psaradakis2005markov} that introduces different terminologies for causality based on Granger’s ideas and provide a set of parametric non-causality constraints in the context of Markov switching VAR models.
In a similar context, \citet{bianchi2019modeling} investigates time-varying systemic risk based on a range of multi-factor asset pricing models and develops a Markov Chain Monte Carlo (MCMC) scheme to infer their model parameters and consequently obtain their corresponding networks.
Another attempt is \citet{bonaccolto2019estimation} in which the authors explore quantile-based methods of Granger causality.
%This work focuses on the linear factor model augmented with network dependence introduced by \citet{billio2018networks} and develops a method to identify causality among quantiles of the time series. 
This method is consistent with \citet{hong2009granger} and \citet{corsi2018measuring} that focus on causality among tail events. 
These methods are suitable for capturing causal relationships that are not in the center of their distributions, or in the mean but they are in the tails of their distributions.
It is important to emphasize that our proposed approach using DI is also capable of capturing such causal relations.

Most of the above aforementioned approaches are developed and tested for small size networks. 
Often, the problem of network identification in high dimensional settings requires more considerations and even its own techniques. 
For instance, \citet{billio2019bayesian} proposes a Bayesian non-parametric Lasso prior (BNP-Lasso) for high-dimensional VAR models that can improve efficiency and accuracy. 
In order to overcome overparametrization and overfitting issues in large VAR models, BNP-Lasso clusters the VAR coefficients into groups and shrinks the coefficients of each group toward a common location.
%Although, the proposed method in this work is suitable for inference in high dimensional VAR models but 
However, this method is limited to linear models with Gaussian innovations.
To overcome this limitation, \citet{kalli2018bayesian} proposes a Bayesian non-parametric approach that allows for nonlinearity in the conditional mean, heteroskedasticity in the conditional variance, and non-Gaussian innovations. However, unlike the BNP-Lasso, it does not allow sparsity in the model.
\citet{petrova2019quasi} proposes yet another non-parametric, quasi-Bayesian likelihood estimation methodology for high dimensional setting with time varying parameters. 
%This method is also non-parametric. 
The work in \citet{iacopini2019bayesian} tackles the curse of dimensionality by a two-stage approach. In the first stage, a spike-and-slab prior distribution is used for each entry of the coefficient matrix which also identifies the interconnection network.
In the second stage, it imposes prior dependence on the coefficients by specifying a Markov process for their random distribution. 
%A side product of this procedure is the interconnection network.
A closely related work is \citet{bernardi2019high} that proposes a shrinkage and selection methodology designed for network inference in high-dimensional settings. It uses a regularized linear regression model with spike-and-slab prior on the parameters.
However, both methods are limited to VAR models.

The rest of the paper is organized as follows. 
Section \ref{Sec1} reviews the notion of Granger causality and formally introduce directed information graphs which is suitable for linear and nonlinear systems. 
In Section \ref{sec:method}, we introduce a novel approach for inferring the Granger-causal network of high dimensional nonlinear systems. 
Finally, in Section \ref{sec:exp}, we apply our method to learn the causal network of both synthetic and real-world dataset. For the real-world experiment, we used the daily stock prices of major US firms.

@@

In this section, we present a statistical approach to learn the causal interconnections in a dynamical systems based on Granger causality \citet{granger}. 
We begin by introducing some notations.
Plain capital letters denote random variables or processes, while lowercase letters denote their realizations. 
Bold letters are used for column vectors, matrices, and tensors and calligraphy letters are used for sets. 
We use $X_{j,t}$ to denote the value of a time series $X_j$ at time $t$ 
%and $X_{j,t_1}^{t_2}$, when $t_1\leq t_2$, to denote $X_j$ from time $t_1$ to $t_2$, namely, $(X_{j,t_1},...,X_{j,t_2})$. 
%Subsequently, 
and $X^{t}_j$ to denote the time series $X_j$ up to time $t$. 
For a set $\mathcal{A}=\{a_1,...,a_n\}$ and an index set $\mathcal{I}\subseteq\{1,...,n\}$, we define $\mathcal{A}_{-\mathcal{I}}:=\mathcal{A}\setminus\{a_i: i\in\mathcal{I}\}$.

@@

Researchers from different fields have developed various frameworks and graphical models to capture and represent interconnections among variables or processes.
One of the most popular and widely used frameworks in economics is the notion of Granger causality.
The basic idea in this framework was originally introduced by Wiener \citet{wiener1956theory}, and later formalized by Granger \citet{granger}. 
The idea is as follows: ``we say that $X$ is causing $Y$ if we are better able to predict the future of $Y$ using all available information than if the information apart from the past of $X$ had been used."

Despite broad philosophical viewpoint of \citet{granger1963economic}, his formulation for practical implementation was done using multivariate autoregressive (MVAR) models and linear regression which has been widely adopted in econometric and other disciplines.
More precisely, in order to identify the influence of $X_t$ on $Y_t$ in a MVAR comprises of three time series $\{X,Y,Z\}$, Granger's idea is to compare the performance of two linear regressions: the first one predicts $Y_t$ given $\{X^{t-1},Y^{t-1},Z^{t-1}\}$ and the second one predicts $Y_t$ given $\{Y^{t-1},Z^{t-1}\}$. Clearly, the performance of the second predictor is bounded by the first predictor. If they have the same performance, then we say $X$ does not Granger cause $Y$. It is important to emphasize that this formulation is only applicable in linear systems.

To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger causality. 
In this work, we introduce and apply directed information (DI) \citet{directed}, an information-theoretical tools to measure interconnections among firms.
DI has been used in many applications to infer causal relationships. For example, \citet{spike} and \citet{di12} used it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} applied to market data.

In order to visualize the inferred interconnections among time series using DI, directed information graphs (DIGs) have been developed \citet{directed}.  
DIGs are a type of graphical models in which nodes represent time series and arrows indicate the direction of causation. We use DIG to represent the causal network among the covered firms.

@@

In the rest of this section, we describe how the DI can capture the interconnections in causal\footnote{In causal systems, given the full past of the system, the present of the processes become statistically independent. In other words, there are no simulations relationships between the time series.} dynamical systems (linear or non-linear) and formally define DIGs.

@@

At time $t$ for an outcome $y_t\in \mathcal{Y}$, the difference between the log-losses of the two predictors compares their performances.  This difference is also called \textit{regret},
\begin{align}
r_t&:=-\log{p(Y_t=y_t|Y^{t-1},Z^{t-1})}-\big(-\log {p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}\big)\\
&=\log \frac{p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}{p(Y_t=y_t|Y^{t-1},Z^{t-1})}.
\end{align}
Note that the regrets are non-negative for all $t$ and all outcomes $y_t$.
The average regret over the time horizon $[1,T]$ is given by 
\begin{align}\label{eq:ave}
   \frac{1}{T}\sum_{t=1}^T \mathbb{E}[r_t],
\end{align}
where the expectation is taken over the joint density function\footnote{For the sake of notational simplicity, we use $p(y^{t},z^{t-1},x^{t-1})$ to denote $p(Y^{t}=y^{t},Z^{t-1}=z^{t-1},X^{t-1}=x^{t-1})$.} of $X$, $Y$, and $Z$, i.e., 
\begin{align}
    \mathbb{E}[r_t]=\int p(y^{t},z^{t-1},x^{t-1})\log \frac{p(y_t|y^{t-1},z^{t-1},x^{t-1})}{p(y_t|y^{t-1},z^{t-1})}dy^tdx^{t-1}dz^{t-1}.
\end{align}
The average regret in \eqref{eq:ave} is called \textit{directed information} (DI) and will be our measure of causation in this work.
This measure is always positive and if it is zero, it indicates that the history of time series $X$ contains no significant information that would help in predicting the future of time series $Y$ given the history of $Y$ and $Z$. 
This definition can be generalized to more than three time series as follows,

@@

A simple way to represent the DIG $G$ of a dynamical system is via the adjacency matrix $\textbf{DIG}=[d_{i,j}]_{m\times m}$ that is defined by
\begin{align}
    d_{j,i} = \left\{
\begin{array}{ll}
      1 & \text{if}\  I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})>0,\\
      0 & \text{otherwise}. \\
\end{array} 
\right. 
\end{align}
Given a DIG $G=(\mathcal{V},\mathcal{E})$, we define the parent set of node $R_j$ denoted by $\mathcal{PA}_j\subset \mathcal{V}$ to be the set of all times series that have direct influences on $R_j$, i.e., $\mathcal{PA}_j:=\{R_k: d_{j,k}=1\}$. Similarly, the children set of node $R_j$ is given by $\mathcal{CH}_j:=\{R_k: d_{k,j}=1\}$. 
Next example demonstrates the DIG of a simple linear system.
\begin{example}\label{r1}
Consider a network of three times series $\{X,Y,Z\}$ with the following linear dynamic,
\begin{small}
\begin{equation}\label{model1}
\begin{pmatrix}
X_t \\
Y_t\\
Z_t
\end{pmatrix}=\begin{pmatrix}
0.5 & 0& 0\\
0.4 & 0.5 & 0\\
0 & -0.2 & 0
\end{pmatrix}\begin{pmatrix}
X_{t-1} \\
Y_{t-1}\\
Z_{t-1}
\end{pmatrix}+\begin{pmatrix}
N_{X_t} \\
N_{Y_t}\\
N_{Z_t}
\end{pmatrix},
\end{equation}
\end{small}
where $N_X$, $N_Y$, and $N_Z$ are three independent stationary Gaussian processes with zero mean and a diagonal covariance matrix (1, 0.9, 1). 
Since the dynamic is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in general setting.}\citet{acc2014}.
\begin{equation}\label{simueq}
I(Z\rightarrow Y||X)=\frac{1}{2T}\sum_{t=1}^{T}\log\frac{|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},X_{t-1}}|}{|\Sigma_{Y_{t-1},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},Y_{t},X_{t-1}}|},
\end{equation}
where $|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}|$ denotes the determinant of the covariance matrix of $\{Y_{t-1},Y_{t},X_{t-1}\}$. Using \eqref{simueq}, we computed the DIs of this system,
\begin{figure}
\hspace{4cm}
\xygraph{ !{<0cm,0cm>;<1.9cm,0cm>:<0cm,1.7cm>::} 
!{(-.7,1) }*+{X}="x"
!{(.4,1.6) }*+{Y}="y"
!{(1.5,1)}*+{Z}="z"
%"x":@(ld,lu)"x"
%"y":@(lu,ru)"y"
"x":"y"
"y":"z"
}
   \mycaption{DIG of the system in \eqref{model1}}{}
\label{fig:example_dig}
\end{figure}
\begin{small}
\begin{align*}
&I(Y\rightarrow X||Z)=0,\ \ I(Z\rightarrow X||Y)=0,\ \ I(Z\rightarrow Y||X)=0,\ \ I(Z\rightarrow Z||X,Y)=0,\\
&I(X\rightarrow Z||Y)=0,\ \ \ \ I(X\rightarrow Y||Z)\approx 0.1, \ \ \ I(Y\rightarrow Z||X)\approx 0.03.
\end{align*}
\end{small}
Figure \ref{fig:example_dig} illustrates the DIG of this system. In this example, $\mathcal{PA}_Z=\{Y\}$ and $\mathcal{CH}_Z=\{\}$.
\end{example}

@@

\begin{remark}

A causal model allows a factorization of the joint density function in some specific ways.
It was shown in \citet{directed} that under a mild assumption, the joint density function of a causal discrete-time dynamical system with DIG $G(\mathcal{R},\mathcal{E})$ can be factorized as follows,
\begin{equation}\label{abo}
p(R_1,...,R_m)=\prod_{i=1}^{m}\prod_{t=1}^T p(R_{i,t}|\mathcal{R}_{\mathcal{PA}_i\cup\{i\}}^{t-1}).
\end{equation}
%where $\mathcal{B}_i$ denotes the parent set of node $R_i$ in graph $G$ and $\mathcal{R}_{B_i\cup\{i\}}^{t-1}$ denotes the time series with indices $B_i\cup\{i\}$ up to time $t-1$.  
Such factorization is called a generative model.

\end{remark}

\subsection{Inferring  DIGs}
Inferring the DIG of a dynamical system requires estimating the DIs between all ordered pairs of time series within that system. More precisely, inferring the DIG of a network of $m$ time series requires computing $m(m-1)$ number of DIs. 
On the other hand, estimating DI  requires estimating all the expectation terms in (\ref{cdif}).
In information theory this expectation is known as \textit{conditional mutual information}\footnote{For more details see \citet{cover2012elements}.}, i.e., 
\begin{align}
    I(R_{j,t};R_{i}^{t-1}|\mathcal{R}^{t-1}_{-\{i,j\}},R^{t-1}_{j}):=\mathbb{E}\left[\log\frac{p(R_{j,t}|\mathcal{R}^{t-1}_{-\{i,j\}},R_{j}^{t-1},R_{i}^{t-1})}{p(R_{j,t}|\mathcal{R}^{t-1}_{-\{i,j\}},R_{j}^{t-1})}\right].
\end{align}
Using this notation, \eqref{cdif} can be written as follows 
\begin{align}\label{estdi}
I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})=\frac{1}{T}\sum_{t=1}^T I(R_{j,t};R_{i}^{t-1}|\mathcal{R}^{t-1}_{-\{i,j\}},R^{t-1}_{j}),
\end{align}
Therefore, parametric and non-parametric estimators for the conditional mutual information can be used to estimate the DIs. 
There are different methods that can be used to estimate the terms in (\ref{estdi}) given i.i.d. samples of the time series such as plug-in empirical estimator and k-nearest neighbor estimator. For an overview of such estimators see the Appendix and the articles in \citet{paninski2003estimation}, \citet{noshad2019scalable}, and \citet{jiao2013universal}.

>>>>

\section{Introduction}\label{S:2.1}

@@

In the great majority of practical applications, Granger-causality has been studied in the context of Vector Autoregressive (VAR) models. 
For instance, \citet{billio2012econometric} propose several measures  based on Granger-causality to capture the connections between the monthly returns of different financial institutions. 
It uses principle component analysis and pairwise Granger-causality tests to identify the causal networks. 
Other related works are \citet{diebold2014network} and \citet{barigozzi2016network} in which the authors propose connectedness measures based on generalized variance decomposition.
However, the measures introduced in these works are again limited to linear systems and they are based on pairwise comparison which as we show in Section  \ref{sec:dig} fails to infer the true causal relationships.
%and they are based on pairwise comparison which as we show in Section \ref{sec:dig}, it fail to infer the true causal relationships.
 %focuses on one particular network structure: the long-run variance decomposition network. Such networks are characterized by the infinite vector moving average and similar to \citet{diebold2014network}, they are limited to linear systems.

\clearpage
Contributions of this chapter are both in network identification literature and in finance.
%This paper adds contribution in the network identification literature as well as in finance literature. 
Our contributions to network identification are as follows:
\begin{itemize}
    \item We use an information-theoretic measure known as directed information (DI) to infer the Granger-causalities among a set of time series. 
This measure is non-parametric (i.e. it does not depend on the underlying model of the dynamics) and it is capable of capturing causal relationships in both linear and nonlinear systems. The output of this approach is a directed graph known as Directed Information Graph (DIG) that visualizes the interconnections among a set of time series such as stock returns.

\item Computing DI has high computational and sample complexity which makes it not suitable for inferring the causal structure of large networks. 
To overcome this problem, we develop a novel approach based on Recurrent Neural Networks (RNNs) that reduces the complexity of evaluating DIs in high-dimensional settings. 
\end{itemize}

@@

Applications of DIG are various in finance. 
In particular, we recommend to use it as a preliminary feature selection of another predictive model. Feature selection is a process often used in machine learning and statistics which consists of keeping only a subset of relevant features, usually to avoid overfitting or to reduce dimensionality. 
For example, \citet{fs1} and \citet{fs2} show that extraneous features are prone to reduce model's performance measures. 
Finance applications of feature selection models are various and include credit scoring, stock market behavior analysis  or even fraud detection (\citet{fs_fraud}).
\citet{tsai2009} states that feature selection preprocessing is not addressed carefully enough in the bankruptcy prediction literature. They compare five feature selection methods used in bankruptcy prediction: t-test, correlation matrix, step-wise regression, Principle Component Analysis (PCA) and factor analysis and show that any of these methods improves performance. 
\citet{fs_stock} uses a Sequential Forward Selection algorithm to select relevant features predicting the Turkish market index. The use of such feature selection model reduces model prediction error compared to the case where all features are used. This is due to information embedded in several economic factors already included in the market index.

@@

\subsection{Related Work}
In recent years, several approaches have been developed to generalize the applicability of Granger-causality to non-linear and large dynamics. 
To mention a few, \citet{psaradakis2005markov} introduce different terminologies for causality based on Granger’s ideas and provide a set of parametric non-causality constraints in the context of Markov switching VAR models.
In a similar context, \citet{bianchi2019modeling} investigate time-varying systemic risk based on a range of multi-factor asset pricing models and develop a Markov Chain Monte Carlo (MCMC) scheme to infer their model parameters and consequently obtain their corresponding networks.
Another attempt is \citet{bonaccolto2019estimation} in which the authors explore quantile-based methods of Granger-causality.
%This work focuses on the linear factor model augmented with network dependence introduced by \citet{billio2018networks} and develops a method to identify causality among quantiles of the time series. 
This method is consistent with \citet{hong2009granger} and \citet{corsi2018measuring} that focus on causality among tail events. 
These methods are suitable for capturing causal relationships that are not in the center of their distributions, or in the mean but in the tails of their distributions.
It is important to emphasize that our proposed approach using DI is also capable of capturing such causal relations.

Most of the above aforementioned approaches are developed and tested for small size networks. 
Often, the problem of network identification in high dimensional settings requires more considerations and even its own techniques. 
For instance, \citet{billio2019bayesian} propose a Bayesian non-parametric Lasso prior (BNP-Lasso) for high-dimensional VAR models that can improve efficiency and accuracy. 
In order to overcome overparametrization and overfitting issues in large VAR models, BNP-Lasso clusters the VAR coefficients into groups and shrinks the coefficients of each group toward a common location.
%Although, the proposed method in this work is suitable for inference in high dimensional VAR models but 
However, this method is limited to linear models with Gaussian innovations.
To overcome this limitation, \citet{kalli2018bayesian} propose a Bayesian non-parametric approach that allows for nonlinearity in the conditional mean, heteroskedasticity in the conditional variance, and non-Gaussian innovations. However, unlike the BNP-Lasso, it does not allow sparsity in the model.
\citet{petrova2019quasi} proposes yet another non-parametric, quasi-Bayesian likelihood estimation methodology for high dimensional setting with time-varying parameters. 
%This method is also non-parametric. 
The work in \citet{iacopini2019bayesian} tackles the curse of dimensionality by a two-stage approach. In the first stage, a spike-and-slab prior distribution is used for each entry of the coefficient matrix which also identifies the interconnection network.
In the second stage, it imposes prior dependence on the coefficients by specifying a Markov process for their random distribution. 
%A side product of this procedure is the interconnection network.
A closely related work is \citet{bernardi2019high} that proposes a shrinkage and selection methodology designed for network inference in high-dimensional settings. It uses a regularized linear regression model with spike-and-slab prior on the parameters.
However, both methods are limited to VAR models.

The rest of the chapter is organized as follows. 
Section \ref{Sec1} reviews the notion of Granger-causality and formally introduces directed information graphs which are suitable for linear and nonlinear systems. 
In Section \ref{sec:method}, we introduce a novel approach for inferring the Granger-causal network of high dimensional nonlinear systems. 
Finally, in Section \ref{sec:exp}, we apply our method to learn the causal network of both synthetic and real-world dataset. For the real-world experiment, we use the daily stock prices of US firms.

@@

In this section, we present a statistical approach to learn the causal interconnections in a dynamical systems based on Granger-causality (see \citet{granger}). 
We begin by introducing some notations.
Plain capital letters denote random variables or processes, while lowercase letters denote their realizations. 
Bold letters are used for column vectors, matrices, and tensors and calligraphy letters are used for sets. 
We use $X_{j,t}$ to denote the value of a time series $X_j$ at time $t$ 
%and $X_{j,t_1}^{t_2}$, when $t_1\leq t_2$, to denote $X_j$ from time $t_1$ to $t_2$, namely, $(X_{j,t_1},...,X_{j,t_2})$. 
%Subsequently, 
and $X^{t}_j$ to denote the time series $X_j$ up to time $t$. 
For a set $\mathcal{A}=\{a_1,...,a_n\}$ and an index set $\mathcal{I}\subseteq\{1,...,n\}$, we define $\mathcal{A}_{-\mathcal{I}}:=\mathcal{A}\setminus\{a_i: i\in\mathcal{I}\}$.

@@

Researchers from different fields have developed various frameworks and graphical models to capture and represent interconnections among variables or processes.
One of the most popular and widely used frameworks in economics is the notion of Granger-causality.
The basic idea in this framework was originally introduced by \citet{wiener1956theory}, and later formalized by \citet{granger}. 
The idea is as follows: ``we say that $X$ is causing $Y$ if we are better able to predict the future of $Y$ using all available information than if the information apart from the past of $X$ had been used."

Despite broad philosophical viewpoint of \citet{granger1963economic}, his formulation for practical implementation was done using multivariate autoregressive (MVAR) models and linear regression which has been widely adopted in econometric and other disciplines.
More precisely, in order to identify the influence of $X_t$ on $Y_t$ in a MVAR comprises of three time series $\{X,Y,Z\}$, Granger's idea is to compare the performance of two linear regressions: the first one predicts $Y_t$ given $\{X^{t-1},Y^{t-1},Z^{t-1}\}$ and the second one predicts $Y_t$ given $\{Y^{t-1},Z^{t-1}\}$. Clearly, the performance of the second predictor is bounded by the first predictor. If they have the same performance, then we say $X$ does not Granger-cause $Y$. It is important to emphasize that this formulation is only applicable in linear systems.

To go beyond linear systems, works such as \citet{directed} and \citet{massey} use information-theoretical measures and generalize Granger-causality. 
In this work, we introduce and apply directed information (see \citet{directed}), an information-theoretical tools to measure interconnections among firms.
Directed information (DI) has been used in many applications to infer causal relationships. For example, \citet{spike} and \citet{di12} use it for analyzing neuroscience data and \citet{acc2014}  and \citet{etesami2018econometric} apply to market data.

Directed information graphs (DIGs) have been developed to visualize the inferred interconnections among time series (see \citet{directed}).  
DIGs are a type of graphical models in which nodes represent time series and arrows indicate the direction of causation. We use DIG to represent the causal network among the covered firms.

@@

In the rest of this section, we describe how the DI can capture the interconnections in causal dynamical systems (linear or non-linear) and formally define DIGs.

@@

At time $t$ for an outcome $y_t\in \mathcal{Y}$, the difference between the log-losses of the two predictors compares their performances.  This difference is also called regret, denoted $r_t$ :
\begin{align}
r_t:&=-\log{p(Y_t=y_t|Y^{t-1},Z^{t-1})}-\big(-\log {p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}\big) \nonumber \\
&=\log \frac{p(Y_t=y_t|Y^{t-1},Z^{t-1},X^{t-1})}{p(Y_t=y_t|Y^{t-1},Z^{t-1})}.
\end{align}
Note that the regrets are non-negative for all $t$ and all outcomes $y_t$.
The average regret over the time horizon $[1,T]$ is given by 
\begin{align}\label{eq:ave}
   \frac{1}{T}\sum_{t=1}^T \mathbb{E}[r_t],
\end{align}
where the expectation is taken over the joint density function\footnote{For the sake of notational simplicity, we use $p(y^{t},z^{t-1},x^{t-1})$ to denote $p(Y^{t}=y^{t},Z^{t-1}=z^{t-1},X^{t-1}=x^{t-1})$.} of $X$, $Y$, and $Z$, i.e., 
\begin{align}
    \mathbb{E}[r_t]=\int p(y^{t},z^{t-1},x^{t-1})\log \frac{p(y_t|y^{t-1},z^{t-1},x^{t-1})}{p(y_t|y^{t-1},z^{t-1})}dy^tdx^{t-1}dz^{t-1}.
\end{align}
The average regret in \eqref{eq:ave} is called directed information (DI) and will be our measure of causation in this work.
This measure is always positive and if it is zero, it indicates that the history of time series $X$ contains no significant information that would help in predicting the future of time series $Y$ given the history of $Y$ and $Z$. 
This definition can be generalized to more than three time series as follows,

@@

A simple way to represent the DIG $G$ of a dynamical system is via the adjacency matrix $\textbf{DIG}=[d_{i,j}]_{m\times m}$ that is defined by
\begin{align}
    d_{j,i} = \left\{
\begin{array}{ll}
      1 & \text{if}\  I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})>0,\\
      0 & \text{otherwise}. \\
\end{array} 
\right. 
\end{align}
Given a DIG $G=(\mathcal{V},\mathcal{E})$, we define the parent set of node $R_j$ denoted by $\mathcal{PA}_j\subset \mathcal{V}$ to be the set of all times series that have direct influences on $R_j$, i.e., $\mathcal{PA}_j:=\{R_k: d_{j,k}=1\}$. Similarly, the children set of node $R_j$ is given by $\mathcal{CH}_j:=\{R_k: d_{k,j}=1\}$. 
Example \ref{r1} demonstrates the DIG of a simple linear system.
\begin{example}\label{r1}
Consider a network of three times series $\{X,Y,Z\}$ with the following linear dynamic,
\begin{small}
\begin{equation}\label{model1}
\begin{pmatrix}
X_t \\
Y_t\\
Z_t
\end{pmatrix}=\begin{pmatrix}
0.5 & 0& 0\\
0.4 & 0.5 & 0\\
0 & -0.2 & 0
\end{pmatrix}\begin{pmatrix}
X_{t-1} \\
Y_{t-1}\\
Z_{t-1}
\end{pmatrix}+\begin{pmatrix}
N_{X_t} \\
N_{Y_t}\\
N_{Z_t}
\end{pmatrix},
\end{equation}
\end{small}
where $N_X$, $N_Y$, and $N_Z$ are three independent stationary Gaussian processes with zero mean and a diagonal covariance matrix (1, 0.9, 1). 
Since the dynamic is linear and the exogenous noises are Gaussian, we can compute the DIs using the following expression\footnote{Equation \eqref{simueq} does not hold in more general settings.}(see \citet{acc2014}).
\begin{equation}\label{simueq}
I(Z\rightarrow Y||X)=\frac{1}{2T}\sum_{t=1}^{T}\log\frac{|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},X_{t-1}}|}{|\Sigma_{Y_{t-1},X_{t-1}}||\Sigma_{Z_{t-1},Y_{t-1},Y_{t},X_{t-1}}|},
\end{equation}
where $|\Sigma_{Y_{t-1},Y_{t},X_{t-1}}|$ denotes the determinant of the covariance matrix of $\{Y_{t-1},Y_{t},X_{t-1}\}$. Using \eqref{simueq}, we compute the DIs of this system,
\begin{small}
\begin{align}  \label{di_xyz}
&I(Y\rightarrow X||Z)=0,\ \ I(Z\rightarrow X||Y)=0,\ \ I(Z\rightarrow Y||X)=0,\ \ I(Z\rightarrow Z||X,Y)=0,\\
&I(X\rightarrow Z||Y)=0,\ \ \ \ I(X\rightarrow Y||Z)\approx 0.1, \ \ \ I(Y\rightarrow Z||X)\approx 0.03. \nonumber
\end{align}
\end{small}
Figure \ref{fig:example_dig} illustrates the DIG of the system \eqref{model1}. In this example, $\mathcal{PA}_Z=\{Y\}$ and $\mathcal{CH}_Z=\{\}$.
\begin{figure}[h]
\hspace{4cm}
\xygraph{ !{<0cm,0cm>;<1.9cm,0cm>:<0cm,1.7cm>::} 
!{(-.7,1) }*+{X}="x"
!{(.4,1.6) }*+{Y}="y"
!{(1.5,1)}*+{Z}="z"
%"x":@(ld,lu)"x"
%"y":@(lu,ru)"y"
"x":"y"
"y":"z"
}
\mycaption{DIG estimated from the directed informations \eqref{di_xyz}}{X influences Y, Y influences Z. This DIG correctly estimates the dynamics of the system in \eqref{model1}.}
\label{fig:example_dig}
\end{figure}
\end{example}

@@

\clearpage

\begin{remark}
A causal model allows a factorization of the joint density function in some specific ways.
It was shown in \citet{directed} that under a mild assumption, the joint density function of a causal discrete-time dynamical system with DIG $G(\mathcal{R},\mathcal{E})$ can be factorized as follows,
\begin{equation}\label{abo}
p(R_1,...,R_m)=\prod_{i=1}^{m}\prod_{t=1}^T p(R_{i,t}|\mathcal{R}_{\mathcal{PA}_i\cup\{i\}}^{t-1}).
\end{equation}
%where $\mathcal{B}_i$ denotes the parent set of node $R_i$ in graph $G$ and $\mathcal{R}_{B_i\cup\{i\}}^{t-1}$ denotes the time series with indices $B_i\cup\{i\}$ up to time $t-1$.  
Such factorization is called a generative model. 
\end{remark}

\subsection{Inferring  DIGs}
Inferring the DIG of a dynamical system requires estimating the DIs between all ordered pairs of time series within that system. More precisely, inferring the DIG of a network of $m$ time series requires computing $m(m-1)$ number of DIs. 
On the other hand, estimating DI  requires estimating all the expectation terms in (\ref{cdif}).
In information theory this expectation is known as \textit{conditional mutual information}\footnote{See \citet{cover2012elements} for more details.}, i.e., 
\begin{align}
    I(R_{j,t};R_{i}^{t-1}|\mathcal{R}^{t-1}_{-\{i,j\}},R^{t-1}_{j}):=\mathbb{E}\left[\log\frac{p(R_{j,t}|\mathcal{R}^{t-1}_{-\{i,j\}},R_{j}^{t-1},R_{i}^{t-1})}{p(R_{j,t}|\mathcal{R}^{t-1}_{-\{i,j\}},R_{j}^{t-1})}\right].
\end{align}
Using this notation, \eqref{cdif} can be written as follows 
\begin{align}\label{estdi}
I(R_i\rightarrow R_j||\mathcal{R}_{-\{i,j\}})=\frac{1}{T}\sum_{t=1}^T I(R_{j,t};R_{i}^{t-1}|\mathcal{R}^{t-1}_{-\{i,j\}},R^{t-1}_{j}),
\end{align}
Therefore, parametric and non-parametric estimators for the conditional mutual information can be used to estimate the DIs. 
There are different methods that can be used to estimate the terms in (\ref{estdi}) given i.i.d. samples of the time series such as plug-in empirical estimator and k-nearest neighbor estimator. For an overview of such estimators see Appendix \ref{appendix2} and the articles in \citet{paninski2003estimation}, \citet{noshad2019scalable}, and \citet{jiao2013universal}.

 In general, estimating the DI in (\ref{estdi}) is a complicated task and has high sample complexity.
 This is due to the fact that it requires estimating high dimensional conditional distributions. 
However, knowing some side information about the underlying dynamic can simplify the learning task of the DIG. 
For instance, in Example \ref{r1}, since the underlying dynamic is linear with Gaussian exogenous noises, the DIs can be computed via the covariance matrices \eqref{simueq}. 
Clearly, the covariance matrix can be estimated with lower complexity compared to conditional mutual information. 
For our experimental results, we used \eqref{simueq} for the linear Gaussian experiment and  the k-nearest method in \citet{sricharan2011k} for the non-linear experiment. 
The main reason for selecting k-nearest method is because it has shown relatively better performance compared to the other non-parametric estimators. 
%More precisely, we estimated each of the above conditional mutual information using the $k$-nearest method of \citet{sricharan2011k}. 
For the sake of completeness, we describe the steps of this method in Appendix \ref{appendix2}.
